Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems

### 2024-06-18 final/oup-authoring-template.tex
Add structured abstract

%\abstract{
%\textbf{Motivation:} .\\
%\textbf{Results:} .\\
%\textbf{Availability:} .\\
%\textbf{Contact:} \href{name@email.com}{name@email.com}\\
%\textbf{Supplementary information:} Supplementary data are available at \textit{Journal Name}
%online.}

\abstract{
  \Acfp{CRN} play a pivotal role in diverse fields
  such as systems biology, biochemistry, chemical engineering, and epidemiology.
  High-level definitions of \acp{CRN} enables to use various simulation approaches,
  including deterministic and stochastic methods,
  from the same model.
  However,
  existing Python tools for simulation of \ac{CRN} typically wrap external C/C++ libraries for
  model definition, translation into equations and/or numerically solving them,
  limiting their extensibility and integration with the broader Python ecosystem.
  In response, we developed Poincaré and SimBio,
  two novel Python packages for simulation of dynamical systems and \acp{CRN}.
  Poincaré serves as a foundation for dynamical systems modelling,
  while SimBio extends this functionality to \acp{CRN},
  including support for the \ac{SBML}.
  Poincaré and SimBio are developed as pure Python packages
  enabling users to easily extend their simulation capabilities
  by writing new or leveraging other Python packages.
  Moreover, this does not compromise the performance,
  as code can be \ac{JIT} compiled with Numba.
  Our benchmark tests using curated models from the BioModels repository
  demonstrate that these tools may provide a potentially
  superior performance advantage compared to other existing tools.
  Additionally, to ensure a user-friendly experience,
  our packages use standard typed modern Python syntax that provides a seamless integration with \acp{IDE}.
  Our Python-centric approach significantly enhances code analysis, error detection, and refactoring capabilities,
  positioning Poincaré and SimBio as valuable tools for the modelling community.
}
\keywords{poincare, SimBio, Python, ODE}
\acresetall % reset acronyms after abstract

>>>>

\abstract{
\textbf{Motivation:}
  \Acfp{CRN} play a pivotal role in diverse fields
  such as systems biology, biochemistry, chemical engineering, and epidemiology.
  High-level definitions of \acp{CRN} enables to use various simulation approaches,
  including deterministic and stochastic methods,
  from the same model.
  However,
  existing Python tools for simulation of \ac{CRN} typically wrap external C/C++ libraries for
  model definition, translation into equations and/or numerically solving them,
  limiting their extensibility and integration with the broader Python ecosystem.
\\ \textbf{Results:}
  In response, we developed Poincaré and SimBio,
  two novel Python packages for simulation of dynamical systems and \acp{CRN}.
  Poincaré serves as a foundation for dynamical systems modelling,
  while SimBio extends this functionality to \acp{CRN},
  including support for the \ac{SBML}.
  Poincaré and SimBio are developed as pure Python packages
  enabling users to easily extend their simulation capabilities
  by writing new or leveraging other Python packages.
  Moreover, this does not compromise the performance,
  as code can be \ac{JIT} compiled with Numba.
  Our benchmark tests using curated models from the BioModels repository
  demonstrate that these tools may provide a potentially
  superior performance advantage compared to other existing tools.
  Additionally, to ensure a user-friendly experience,
  our packages use standard typed modern Python syntax that provides a seamless integration with \acp{IDE}.
  Our Python-centric approach significantly enhances code analysis, error detection, and refactoring capabilities,
  positioning Poincaré and SimBio as valuable tools for the modelling community.
\\ \textbf{Availability:}
  Poincaré and SimBio are released under the MIT license.
  their source code is available on GitHub
  (\url{https://github.com/maurosilber/poincare}
  and \url{https://github.com/hgrecco/simbio}).
  and can be installed from PyPI or conda-forge.
\\ \textbf{Contact:} \href{email:hgrecco@df.uba.ar}{hgrecco@df.uba.ar}
}

\keywords{poincare, SimBio, Python, ODE}
\acresetall % reset acronyms after abstract
